% Content: Classical Mechanics
%------------------------------

\section{Classical Mechanics (20\%)}

It includes kinematics, Newton's laws, work and energy, oscillatory motion, rotational motion about a fixed axis, dynamics of systems of particles, central forces and celestial mechanics, three-dimensional particle dynamics, Lagrangian and Hamiltonian formalism, non-inertial reference frames, and elementary topics in fluid dynamics.

\subsection{Kinematics}

\subsubsection{Linear Motion in the inertial frame of reference}

\subsubsection*{Average velocity}
\begin{align*}
v = \frac{\Delta x}{\Delta t} = \frac{x_{2} - x_{1}}{t_{2} - t_{1}}
\end{align*}

\subsubsection*{Instantaneous velocity}
\begin{align*}
v = \lim_{\Delta t \to 0} \frac{\Delta x}{\Delta t} = \frac{dx}{dt} = v(t)
\end{align*}

\subsubsection*{Acceleration}
\begin{align*}
a = \lim_{\Delta t \to 0} \frac{\Delta v}{\Delta t} = \frac{dv}{dt} = a(t)
\end{align*}

If acceleration is zero then, the velocity will be constant. Likewise, if velocity is constant then, acceleration will be zero.

\subsubsection*{Kinematic equations of motion under constant acceleration: SUVAT rule}
\begin{align}
v &= u + at\\
(x - x_{0}) = s &= ut + \frac{1}{2}at^{2} \\
v^{2} &= u^{2} + 2as \\
(x - x_{0}) &= \frac{1}{2}(u + v)t 
\end{align}

\subsubsection{Projectile motion}

To solve projectile motion related problems, we only need equations of motion for \emph{two dimensions}. This is because if you observe an object moving in the projectile motion, you will only see two components to study this motion. But I want to aware you this is not true for an object which is moving in a curvy way (or, approximately zigzag or spiral). For example, boomerang. So, for the projectile motion, we only need is

\begin{align*}
\text{in $x$-axis:\quad} x(t_{1}) &= x(t_{0}) + \mathcolorbox{cyan}{u_{x}}t_{1} \\
\text{, and in $y$-axis:\quad} y(t_{1}) &= y(t_{0}) + \mathcolorbox{cyan}{u_{y}}t_{1} \mathcolorbox{lightgray}{ - \frac{1}{2}g t^{2}_{1}}
\end{align*}
where \hlcyan{$u_{i}$ refers to initial velocity in $i$-axis}. In this type of motion, we assume an object will only falling under the influence of gravity so, we have additional term which is highlighted by gray color. In order to have mathematically correct equation, we also need negative sign in-front of it which suggest us, we need to change the direction of gravity along with the $y$-axis direction. 

\subsubsection{Uniform circular motion}

The most simplest form of an orbit is circular path in-which its acceleration vector is decomposed into radial and tangential components. If its tangential acceleration is zero which means its tangential velocity is constant then, we can have uniform circular motion about the center of circle.  This means if you see tangential velocity be constant then, you can think of the problem is related to circular motion. There must be radially inward force acting on the object which we called as centripetal force. Thus, it keeps an object moving in a circular path with constant speed\footnote{We use ``speed" instead of velocity because we consider our velocity to be constant that means we don't need to deal with the direction of this velocity. Because whatever the direction is, its magnitude of the velocity remains same. But note it, its magnitude play a very crucial role to be in a circular path.}. But, keep in mind that its radial acceleration is non-zero.

\subsubsection*{Radial or centripetal acceleration}
\begin{align*}
a = \frac{v^{2}}{r}
\end{align*}

\subsubsection*{Angle}
\begin{align*}
\theta = \frac{\text{arc}}{\text{radius}}
\end{align*}

\subsubsection*{Angular velocity}
\begin{align*}
\omega = \lim_{\Delta t \to 0}\frac{\Delta \theta}{\Delta t}
\end{align*}

\subsubsection*{Relation of v and $\omega$}
\begin{align}
v = r\omega
\end{align}

\subsubsection*{Relation of a and $\omega$}
\begin{align}
a = r\omega^{2}
\end{align}

\subsubsection*{Angular acceleration}
\begin{align*}
\alpha = \lim_{\Delta t \to 0}\frac{\Delta \omega}{\Delta t}
\end{align*}

\subsubsection*{Rotational equation of motion under constant angular acceleration}
\begin{align}
\omega &= \omega_{0} + \alpha t \\
\theta &= \omega_{0}t + \frac{1}{2}\alpha t^{2} \\
\omega^{2} &= \omega_{0}^{2} + 2\alpha \theta \\
\theta &= \frac{1}{2}(\omega_{0} + \omega)t
\end{align}

\subsubsection*{Centripetal or radial force}
\begin{align}
F_{\text{radial}} = \frac{mv^{2}}{r}
\end{align}

If you see any problem related to circular motion then, any force (i.e. gravitational force or magnetic force or tension on the string) can be equal to centripetal force. 
\begin{align*}
\text{Any force} = F_{\text{radial}}
\end{align*}
For example: In an uniform magnetic field which is perpendicular to the direction of motion of a particle then, its tangential velocity will be constant and can be considered the problem of circular motion.


\subsection{Newton's laws of motion}

\noindent\textbf{First law:} A body continues in its state of rest or \emph{uniform motion} unless any external \emph{unbalanced force} act upon it.
\\
\textbf{Second law:} The net force \emph{on} a body is proportional to its rate of change of \emph{linear momentum}.
\begin{align*}
F = \frac{dp}{dt} = ma
\end{align*}
\\
\textbf{Third law:} When a particle A exerts a force on another particle B (i.e. $F_{AB}$), B simultaneously exerts a force on A (i.e. $F_{BA}$) with the \emph{same magnitude} in the \emph{exactly opposite direction}.
\begin{align*}
F_{AB} = - F_{BA}
\end{align*}

\subsection{Frictional Force}

\begin{align}
F_{f} = \mu \times F_{N}
\end{align}

where, $F_{f}$ refers to frictional force, $F_{N}$ refers to normal force, and $\mu$ refers to coefficient of the friction.

\subsubsection*{Tips}

\begin{itemize}
\item When interactions between the blocks become important, for example when they exert forces on one another through friction, then we must usually treat them as independent objects.

\item If one block of mass ``m" is sits on the top of the another block of mass ``M", and the blocks don't slip then, two bodies are behaving as one with mass of (m+M).  This means, $\mu$ depends on the combination of ``m" and ``M", rather than ``m" or ``M" individually. Also, both the blocks will experience the same acceleration. This makes our problem-solving time faster because we will only use the above concepts and Newton's second law. But not the free body diagram procedure which is actually time-consuming.

\item We change physical quantities in the form of components consists of $sin$ and $cos$. So, we need to remember that at the angle of $45^{\circ}$, $sin(45^{\circ}) = cos(45^{\circ})$ which will make our calculation much simpler.

\item If a spherical or cylindrical object move by rolling (not by slipping) then, frictional force is responsible for it. So, we can neglect translation kinetic energy, and only use rotational kinetic energy. Remember, frictional force only causes rotational motion and which is instantaneous, so you can say frictional force has done negligible work thus, you don't have to account the frictional energy (i.e. $F_{f} \times \text{distance}$).


\end{itemize}

\subsection{Energy}

The most important physical quantity called energy has its own conservation law which helps to solve many problems related to it. And this says: ``\emph{Energy neither be created nor be destroyed, but can be transform from one form to another}". For example: kinetic to potential.

\subsubsection{Energy conservation for conservative force} \label{ecf}

If an object is acted on only by \emph{conservative} forces (i.e. the work done by the force is independent of path. For e.g. gravitational, magnetic, electrical, spring force and etc) then, the sum of the kinetic and potential energies always remains \emph{constant} along the object's path. i.e. E$_\text{initial}$ = E$_\text{final}$. You can use this concept in the problems like elastic collision.

\subsubsection{Energy conservation for non-conservative force}

For non-conservative force, \hlgray{E$_\text{initial}$ $\neq$ E$_\text{final}$}. So, we need to add extra term in which energy is converted into another form. i.e. E$_\text{initial}$ + W$_\text{other}$ = E$_\text{final}$ where W$_\text{other}$ is the work done by non-conservative force. This is we called as \emph{work-energy theorem}. The most common example of the non-conservative force is frictional force. You can use this concept in the problems like in-elastic collision.

\subsubsection*{Tips}
\begin{itemize}
\item Use \emph{conservation of energy}, when we need to calculate how fast (i.e. velocity ?) and how far (i.e. position ?) something goes. Because this law only uses kinetic energy (which depend on velocity), and potential energy (which depend on positions).
\item Use \emph{Kinematics}, when we need to calculate how much time something takes. Because the position and velocity is explicitly a function of time.
\end{itemize}
So, the right choice from these can helps us to solve the problem very fast.

\subsubsection{Types of energies}
The most common types of energies are:
\begin{align}
\text{Translation kinetic energy \hlgray{(T.KE)}} &= \frac{1}{2}m v^{2} \\
\text{Rotational kinetic energy \hlgray{(R.KE)}} &= \frac{1}{2} I \omega^{2} \quad \text{, where $I$ means moment of inertia}\\
\text{Kinetic energy for a body }& \nonumber\\
\text{about its center of mass} &= \text{\hlgray{(T.KE)}} + \text{\hlgray{(R.KE)}} \\
\text{Gravitational potential energy} &= mgh \quad \text{, always use value of $g$ to be 10 $m/s^{2}$} \\
\text{Spring potential energy} &= \frac{1}{2}k x^{2} \quad \text{, where $k$ means spring constant}
\end{align}
So, the simplest trick you can use while doing most of the problems is:
\begin{align*}
\text{Other forms of energy} = \text{Potential energy}
\end{align*}

\subsection{Momentum}
\begin{align*}
\text{Linear momentum: } p &= mv \\
\text{Angular momentum: } L &= I\omega
\end{align*}

\subsubsection{Relation between angular and linear momentum}
\begin{align}
L = r \times p
\end{align}
where $\times$ is the cross-product. This expression is only used when we need to account the system which is build out of smaller pieces. So, you can simply use $ L = I\omega$ for the compact/dense system.

\subsubsection{Conservation of Momentum}
Like conservation of energy, momentum is always conserved in a system in the absence of an external forces. i.e.
\begin{align*}
\text{Total initial momentum} = \text{Total final momentum}
\end{align*}


Always give conservation of momentum as a first priority. Because you cannot use conservation of energy \ref{ecf} for in-elastic collision but, you can still use conservation of momentum. 

\subsection{Torque}

It is also called as moment or moment of force or rotational force.

\begin{align}
\tau &= \frac{dL}{dt} \nonumber \\
\tau &= r \times F 
\end{align}

\subsection{Moment of Inertia}
It is also called as rotational inertia or angular mass. So, moment of inertia can be analogous to an inertial mass\footnote{Inertial mass measures an object's resistance to being linearly accelerated by a force (represented by the relationship $F = ma$). Similarly, angular mass measures an object's resistance to being angularly accelerated by a rotational force.} (or simply mass).

\subsubsection*{A point particle of mass m:}
\begin{align*}
I = m r^{2}
\end{align*}
\subsubsection*{A system of many particles:}
Its moment of inertia is the sum of the individual moments of inertia. But, for an object with infinitely many particles:
\begin{align*}
I = \int r^{2} dm
\end{align*}
where, $dm$ is an infinitesimal mass element and $r$ is the distance from the point to the axis of rotation.

\subsubsection{Parallel axis theorem}
It says, if you know the moment of inertia of a system of mass $M$ rotating about an axis through its center of mass (CM), then its moment of inertia about \emph{any axis parallel} to the CM axis is:
\begin{align}
I = I_{\text{CM}} + M r^{2}.
\end{align}
In pGRE formula sheets, the test taker provides moment of inertia for solid objects like rod, disc, and sphere. Remember, \emph{the moment of inertia about the \emph{CM} axis for a solid cylinder is same as of disc which is not given in the sheet.} i.e. $\frac{1}{2}M r^{2}$.

\subsection{Center of Mass}
It is a point (like a fulcrum) which balances the two sides.

\subsubsection*{A system of point masses}
\begin{align*}
r_{\text{CM}} = \frac{\sum_{i} r_{i} m_{i}}{\sum_{i} m_{i}}
\end{align*}

\subsubsection*{A system of bulk mass}
\begin{align*}
r_{\text{CM}} = \frac{\int r dm}{\int dm}
\end{align*}
 
\subsection*{Confused words in the Classical Mechanics}

\begin{itemize}
\item \emph{String} (i.e. straight wire or thread) and \emph{Spring} (i.e. a helical metal coil).
\item \emph{Gravity} (i.e. gives the answer why every objects fall towards the earth) and \emph{Cavity} (i.e. a hollow space inside solid or liquid)
\item \emph{Geostationary orbit} and \emph{Geosynchronous orbit} are just a synonym of each others. 
\item \emph{Rotation} (i.e. rotate around its own axis) and \emph{Revolution} (i.e. rotate around the heaver object)
\end{itemize}